Analysis of observational data, to calculate the Delayed-Tau model parameters

### 2025-03-23 thesis.qmd
added the parameter results + plot labels

![Alt text](NR/tau_A_double_plot.png){#fig-A-tau}

@@

![](/NR/A-M_*-c_x.png){#fig-A-M}

Since convergence can not occur for $x\ll 1$, this method should not be used for galaxies in the begging phases of their star formation [^10]. If the assumption, $t_{sf}=13.6$ Gyr for all the galaxies is correct, then an inconstancy in the plot $A_{del}-\text{sSFR}$ should only be observed, for higher values of the sSFR, since, according to the delayed- $\tau$ model, higher sSFR values indicate a rising SFR, while lower SFR values indicate that the bulk of $M_*$ is already produced and the galaxy is in the declining SFR phase. However, as seen from @fig-A-sSFR, this inconsistency occurs in galaxies for both high and low sSFR's, meaning that the the model has convergence issues both for slow rising SFR and quenching galaxies, indicating an intrinsic problem with this approach.

@@

Using these priors, the metrics of the Stan programm give good results ($\hat{R}<1.001$ and $\text{ESS} >1000$), while the calculated values correspond to the data of the catalog. Specifically, we use the data to recalculate the SFR from the predicted parameters, using the delayed-$\tau$ model and compare it to the observational data (@fig-comp-sfr-obs-pred), and also compare the normalization constant to the stellar masses of the galaxies (@fig-A-M-mcmc).

::: {#fig-comp-sfr-obs-pred layout-ncol="2"}
![MCMC run for $\tau \sim \text{uniform}$ prior.](r_mcmc_uni/plots/sfr_diff_plot.png){#fig-comp-sfr-obs-pred-uni}

![MCMC run $\tau \sim \text{normal}$ prior.](r_mcmc_normal/plots/sfr_diff_plot.png){#fig-comp-sfr-obs-pred-norm}

Comparison of the observed SFR with the predicted by the parameters given by the MCMC run.
:::

::: {#fig-A-M-mcmc layout-ncol="2"}
![Uniform $\tau$ prior](method_comparison/A_M_MCMC_uni.png){#fig-A-M-mcmc-uni}

@@

Linear regression for $A_{del}-M_*$ for the $A_{del}$ values from the two MCMC runs. The black dashed line represents the regresion weighted by $x$. For the left plot, the regression yields a correlation of 94%, while the weighted regression is the fit $\log(A_{del}) = 0.95\cdot\log(M_*)+0.60$, with a correlation of 98%. For the right plot, the correlation is higher, at 99%, with a weighted regression of $\log(A_{del}) = 0.97\cdot\log(M_*)+0.44$, with a correlation of 99.5%.
:::

>>>>

![Paramater distribution of the delayed-$\tau$ SFH for the galaxies of the LCV. The colourbar represents the present-day stellar mass of the galaxies. The left plot shows the whole parameter range of the model, including negative timescales and $A_{del}$ values that occur by instabilities. In the right plot, the region $\tau \in [0,16]$ and $A_{del}\in[0,10^{13}]$ is depicted. The majority of the galaxies have a timescale of about 3.5-4.5 Gyr as expected from @speagleHighlyConsistentFramework2014](NR/tau_A_double_plot.png){#fig-A-tau}

@@

Since convergence can not occur for $x\ll 1$, this method should not be used for galaxies in the begging phases of their star formation [^10]. If the assumption, $t_{sf}=13.6$ Gyr for all the galaxies is correct, then an inconstancy in the plot $A_{del}-\text{sSFR}$ should only be observed, for higher values of the sSFR, since, according to the delayed- $\tau$ model, higher sSFR values indicate a rising SFR, while lower SFR values indicate that the bulk of $M_*$ is already produced and the galaxy is in the declining SFR phase. However, as seen from @fig-sSFR-A, this inconsistency occurs in galaxies for both high and low sSFR's, meaning that the the model has convergence issues both for slow rising SFR and quenching galaxies, indicating an intrinsic problem with this approach.

@@

Using these priors, the metrics of the Stan programm give good results ($\hat{R}<1.001$ and $\text{ESS} >1000$), while the calculated values correspond to the data of the catalog. Specifically, we use the data to recalculate the SFR from the predicted parameters, using the delayed-$\tau$ model and compare the normalization constant to the stellar masses of the galaxies (@fig-A-M-mcmc).

::: {#fig-A-M-mcmc layout="[[50,50],[70]]"}
![Uniform $\tau$ prior](method_comparison/A_M_MCMC_uni.png){#fig-A-M-mcmc-uni}

@@

![Newton-Raphson Method](/NR/A-M_*-c_x.png){#fig-A-M}

Linear regression for $A_{del}-M_*$ for the $A_{del}$ values from the two MCMC runs and the Newton-Raphson method, with the colorbar repressenting the ratio $x=t_{sf}/\tau$. For the MCMC runs, the black dashed line represents the regresion weighted by $x$. For the upper left plot, the regression yields a correlation of 94%, while the weighted regression is the fit $\log(A_{del}) = 0.95\cdot\log(M_*)+0.60$, with a correlation of 98%. For the upper right plot, the correlation is higher, at 99%, with a weighted regression of $\log(A_{del}) = 0.97\cdot\log(M_*)+0.44$, with a correlation of 99.5%. For the N-R plot, we do two different regressions for the two distinct regions of $A_{del}$, with the upper region only having a correlation of 55\%, while the lower region closely relating the normalization constant with the Stellar Masses of the galaxies, as described in the MCMC plots.
:::
